On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-05-09 Manuscript/manuscript.tex
5/9 Manuscript Udate

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
\eds{a different abbrevation/notation was used for this in the introduction}
\mc{used $BC_a$}
\jy{Isn't this just the empirical CDF? We just need the sample quantile functon.
  What is your reference for this description?}
Let $\hat{G}(c) = \#\{\hat\theta_n^{(b)} < c\}/B$ for $b \in \{1, \ldots, B\}$, let
$z_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$ for $b \in \{1, \ldots, B\}$, let
$\hat{\theta}_{n, -i}$ be the statistic of the sample without the $i^{th}$
block, and let $a$ be $SKEW\{\hat{\theta}_{n, -i}\} / 6$. Then
 $\hat\theta_{BC_a}[\alpha] = \hat{G}^{-1}\Phi(z_0 + \frac{z_{0} + z_{\alpha}}{1 - a(z_{0} + z_{\alpha})}).$ \eds{notation will need to be cleaned and clarified}
A $(1 - \alpha)$ centered $BC_a$ CI is
\[
(\hat\theta_{BC_a}[\alpha/2], \quad
\hat\theta_{BC_a}[1 - \alpha/2]).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. 
\jy{why is it called BCA?}
\mc{addressed}

\paragraph{Standard Normal CI}

\paragraph{Student's t CI}

@@

If $\theta$ is not covered at that $n$, a larger $n$ may be
necessary, assuming that performance will improve as $n$ is increased. $n =$
\jy{learn from how I presented this}
100, 200, 300, 400, and 500, 600, 700 were used in this study. Remember that
the goal of this study is to find the smallest $n$ which is sufficient for
proper coverage.

@@

Based on Figure~\ref{fig:mu}, the performance is
about the same for the three interval types studied for $\mu$. As dependence 
gets stronger,
a larger sample size is necessary. For $\mu$, all three types are working asymptotically 
but for finite sizes, they more or less need about the same sample size to have 
the correct coverage. When $\phi < 0$,
$n = 100$ seems to  be large enough to cover $\mu$ at the nominal rate. When
$\phi = 0$, or in other other words, the sample is independently normally
distributed, %from a normal distribution,
$n = 500$ \eds{$n=400$?} seems to be large enough to cover $\mu$ at the nominal 
rate. Samples
from processes with $\phi > 0$ appear to require a larger $n$ than samples from
processes with $\phi < 0$. When $\phi = 0.2$, $n = 300$ seemed to be necessary
for adequate performance, and when $\phi = 0.4$, an even larger $n$ appeared
to be necessary to properly cover $\mu$. It is rare for $\phi$ to be so high,
so in most cases, if $\phi$ is unknown, a sample of $n \geq 500$ should be
used when the target parameter is $\mu$. \eds{I would argue that none of the 
sample sizes considered work well for $\phi=0.2$ or $\phi=0.4$ based on 
Figure~\ref{fig:mu} and your criteria above (``If all values in the interval are
 below 0.95, the results would suggest that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both.")}

Based on Figure~\ref{fig:sigma}, coverage is
notably lower for $BC_a$ intervals \eds{use consistent notation for this}, 
but $\hat{\theta}_{n}$-centered intervals
and basic intervals have similar performance. For these two interval types,
$n = 500$ seems to be necessary to cover $\sigma$ 
\eds{use $\sigma_x$ instead of $\sigma$} at the nominal rate when
$\phi < 0$. If $\phi = 0$, $n = 400$ seems to large enough to cover $\sigma$ at
the nominal rate. If $\phi = 0.2$, $n = 300$ seems large enough, but if the
process has a greater $\phi$ of $0.4$, $n = 400$ may be necessary. 
\eds{$n=400$ does not appear to be enough to me. I would also argue larger $n$
is needed for the other values of $\phi$ unless the curves asymptote at a value
lower than 0.95. Do the point estimates eventually get closer to the nominal 
level with larger $n$?}  
It is more
common for $\phi$ to be $> 0$ in applied analysis, so in most cases, if $\phi$
is unknown, using a sample of $n \geq 400$ should be sufficient for estimation
of $\sigma$. Based on our results, $BC_a$ intervals should not be used to 
estimate $\sigma$.

When block bootstrap is used to estimate $\phi$ using $BC_a$ and percentile
intervals, the assumption that coverage rates asymptotically approach the
nominal rate as $n$ increases appears to be invalid based on 
Figure~\ref{fig:sigma}.
\eds{Figure~\ref{fig:phi}?}
The $BC_a$ interval performs very poorly and deteriorates more noticeably
when the absolute value $\phi$ gets
higher. Even when $\phi = 0$, $\phi$ is still undercovered for $n$ as high
as $700$. Like the $BC_a$ interval, the percentile interval deteriorates in
coverage when $\phi$ is high, although the coverage is not as low.
Interestingly, when $\phi = 0$, the percentile interval appears to actually
over-cover $\phi$. Regardless of whether $\phi$ is positive or negative, these 
two intervals do not perform well. These results suggest that $BC_a$ and basic 
percentile
block bootstrap intervals should not be used estimate temporal dependence
factors of time series. On the other hand, $\hat{\theta}_{n}$-centered
intervals seem to recover $\phi$ correctly as long as $n \geq 200$, and this
appears to be true regardless of $\phi$, although a greater $\phi$ may result
in a decline in performance.

\jy{One summary paragraph at the end.}

To summarize, the CI type chosen should be based on the target parameter.
For estimating $\mu$, any of the three interval types can be used, but for
$\sigma$, only $\hat{\theta}_{n}$-centered and basic percentile intervals
should be used, and only $\hat{\theta}_{n}$-centered intervals should be used
for $\phi$. When estimating
$\mu$, $n = 500$ should be large enough in situations where $\phi$ is unknown.
When estimating $\sigma$, $n = 400$ is sufficient is situations where $\phi$ is
unknown. If $\phi$ is known, a lesser $n$ may be adequate. When estimating
$\phi$, $n = 200$ is large enough when using the correct type of interval.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_mu}
  \caption{Plots of $\mu$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results from $n$
    ranging from 100-700 were recorded.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_sigma}
  \caption{Plots of $\sigma$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results from $n$
    ranging from 100-700 were recorded.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_phi}
  \caption{Plots of $\phi$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results from $n$
    ranging from 100-700 were recorded.}
  \label{fig:phi}
\end{figure}

As shown in the above figures, only $\hat{\theta}_{n}$-centered intervals
managed to solve the deterioration problem with respect to recovering $\phi$,
indicating that this interval should be used as opposed to the uncentered
percentile interval or the more complex $BC_a$ interval when estimating
$\phi$. Increasing $n$ will not fix the performance of either of these
intervals because the assumption that the coverage rate will approach an upper
limit of .95 as $n$ increases is invalid. As expected for
$\hat{\theta}_{n}$-centered intervals, a higher positive $\phi$ demands a
larger $n$ to estimate all parameters, but a negative $\phi$ may require a
lesser $n$. For example, for both $\phi = 0.2$ and $0.4$, $n = 100$ was
sufficient to recover $\mu$ at a rate of approximately 95\%.
\eds{not sure if above paragraph is needed - seems to repeat information in the 
previous two paragraphs}

@@

When using $\hat{\theta}_{n}$-centered intervals and $\phi$ is
unknown, the results of this study suggest that an $n$ of around 500 may be
necessary for common practice to estimate a simple parameter like $\mu$, and
an $n$ of around 400 may be necessary to
estimate $\sigma$. If $\phi$ is already known, a lesser $n$ may be adequate to
estimate these parameters. When $\phi$ was negative, a $n$ of around 100 was
sufficient to estimate $\mu$. However, in real world applications, $\phi$ is
more commonly found to be positive, so a larger $n$ is more likely to be
necessary. Lastly, to estimate $\phi$, an $n$ of around 400 may be necessary.
Further investigation may be necessary to see if there are other interval
corrections that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods.
This information could also prove to be useful for research using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap interval
construction that would more appropriately recover the parameters of a time
series. \eds{are there other methods that you know of?} 
One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as a moving average process.

>>>>

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
\eds{a different abbrevation/notation was used for this in the introduction}
\mc{used $BC_a$}
\jy{Isn't this just the empirical CDF? We just need the sample quantile functon.
  What is your reference for this description?}
Let $\hat{z}_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$ for $b \in \{1, \ldots, B\}$. 
Let $Z_{(i)}$ be the original sample without the $i$th block $z_i$, and let
$\hat{\theta}_{n, -i}$ be the jackknife values of $Z_{(i)}$ for $i \in \{1, \ldots, n\}$, 
and let $a$ be $SKEW\{\hat{\theta}_{n, -i}\} / 6$. Then
 $\hat\theta_{BC_a}[\alpha] = \hat{G}^{-1}\Phi(z_0 + \frac{z_{0} + z_{\alpha}}{1 - a(z_{0} + z_{\alpha})}).$ \eds{notation will need to be cleaned and clarified}
A $(1 - \alpha)$ centered $BC_a$ CI is
\[
(\hat\theta_{BC_a}[\alpha/2], \quad
\hat\theta_{BC_a}[1 - \alpha/2]).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. 
\jy{why is it called BCA?}
\mc{addressed}

\paragraph{Standard Normal CI}
A $(1 - \alpha$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}*\hat{se}, \hat{\theta}_{n} - z_{(\alpha/2)}*\hat{se}.
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.

\paragraph{Student's t CI}
A $(1 - \alpha$ student's t CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}*\hat{se}, \hat{\theta}_{n} - t_{(\alpha/2), k -1}*\hat{se}.
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.

@@

If $\theta$ is not covered at that $n$, a larger $n$ may be
necessary, assuming that performance will improve as $n$ is increased. $n =$
\jy{learn from how I presented this}
100, 200, 400, 800, and 1600 were used in this study. Remember that
the goal of this study is to find the smallest $n$ which is sufficient for
proper coverage.

@@

% mu
%As dependence gets stronger, a larger sample size is necessary. 
%For $\mu$, Percentile, Centered, and BCA, and Standard more or less 
% need about the same 
% For phi = -0.4, they slightly over-cover mu for all sample sizes, but 
% the error bars generally hover around the 95% line
% For phi = -0.2, all four cover mu at the nominal rate when n = 400, and when 
% n = 1600, although there is a dip in coverage for n = 800.
% Besides that, the coverage rates remain about .9375 for all sample sizes
% For phi = 0, n = 1600 seems to be enough
% For phi = 0.2, the coverage rates are above .9375 when n = 800 and 1600.
% Although there is a slight downward trend from 800 to 1600
% For phi = 0.4, the coverage rates are above .925 when n = 800 and 1600.
% Although there is a more noticeable downward trend from 800 to 1600.
% Student's t performs noticeably better
% For phi = -0.4, it over-covers mu for all sample sizes, and the lower bounds
% are all above the line
%For phi = -0.2, it slightly over-covers mu, and the error bars generally hover
% around the line
% For phi = 0, the interval for mu's coverage rate contains .95 for all sample sizes
% For phi = 0.2 and 0.4, the trend in coverage is similar to the other intervals, but 
% the overall coverage rates are higher.
% The deviations in trend
% for all intervals can be most likely be attributed to noise.

% sigma
%As dependence gets stronger, a larger sample size is necessary. 
% BCA greatly undercovers for all sample sizes, but no trend of coverage deterioration
% Should not be used
%For $\sigma$, Percentile, Centered, and Standard Ci perform similarly
% For phi = -0.4, -0.2, and 0, n = 1600 seems nearly enough to cover sigma at the 
% nomal rate
% For phi = 0.2 and 0.4, a larger sample size seems to be necessary to cover at nominal rate, 
% 0.2: n = 800 and 1600, coverage is above .93 for all three intervals
% 0.4: But at n = 1600, coverage is above .93 for all three intervals
% Student's t performs well
% For phi = -0.4, sigma is covered at the nominal rate for n >= 1600
% For phi = -0.2, sigma is covered at the nominal rate for n >= 400
% For phi = 0, sigma is covered at the nominal rate for n >= 400
% For phi = 0.2, coverage seems to level out at n = 400, but seems to be around .945
% For phi = 0.4, coverage does not seem to increase much from n = 400 onwards, but is above .93
% The deviations in trend
% for all intervlas can be most likely be attributed to noise.

%phi
%As dependence gets stronger, a larger sample size is necessary. 
%Percentile and BCA deteriorate clearly for any level of temporal correlation 
%Regardless of negative or positive
%Should not be used
%Centered and Standard perform similarly
%For these a sample size of n = 800 appears to be necessary to recover phi at the nominal level
% Student's t appears to recover phi at the nominal level for n >= 100
\jy{One summary paragraph at the end.}

To summarize, the CI type chosen should be based on the target parameter.
...

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_mu}
  \caption{Plots of $\mu$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600$ were recorded.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_sigma}
  \caption{Plots of $\sigma_x$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600$ were recorded.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_phi}
  \caption{Plots of $\phi$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600$ were recorded.}
  \label{fig:phi}
\end{figure}

@@

When using $\hat{\theta}_{n}$-centered intervals and $\phi$ is
unknown, the results of this study suggest that an $n$ of around 500 may be
necessary for common practice to estimate a simple parameter like $\mu$, and
an $n$ of around 400 may be necessary to
estimate $\sigma_x$. If $\phi$ is already known, a lesser $n$ may be adequate to
estimate these parameters. When $\phi$ was negative, a $n$ of around 100 was
sufficient to estimate $\mu$. However, in real world applications, $\phi$ is
more commonly found to be positive, so a larger $n$ is more likely to be
necessary. Lastly, to estimate $\phi$, an $n$ of around 400 may be necessary.
Further investigation may be necessary to see if there are other interval
corrections that fix the coverage deterioration problem for $\phi$.

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods.
This information could also prove to be useful for research using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap interval
construction such as $ABC$ or bootstrap-t intervals
that would more appropriately recover the parameters of a time
series. \eds{are there other methods that you know of?} 
One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as a moving average process.
